Modeling the impacts of replacing car trips with combined public transport and cycling: Reproducible methods, results and actionable evidence from biclaR

### 2023-11-30 paper/PaperCEUS/PaperCEUS.tex
final paper to submit

Emission factors are estimated for the following air pollutants: CO,
NO\textsubscript{X}, VOC, and PM. Emission factors of the main
greenhouse gases (GHG) are also estimated: CO\textsubscript{2},
CH\textsubscript{4} and N\textsubscript{2}O, converted in
CO\textsubscript{2}eq by the following relationship\footnote{The weights
  correspond to the Global Warming Potentials (GWP) defined for a
  100-year period by the IPCC in its
  \href{https://www.ipcc.ch/report/ar5/}{5th Assessment Report}.}:
\(EF_{CO_2eq} = EF_{CO_2} + 28\cdot EF_{CH_4} + 265\cdot EF_{N_2O}\).
The CH\textsubscript{4} and N\textsubscript{2}O emission factors do not
vary with travel speed. The PM10 emission factor does not vary with trip
distance.

@@

The conversion of avoided emissions into avoided welfare loss and
respective monetary valuation was based on the EU Guide to Cost-benefit
Analysis (Sartori et al., 2014) and the best up-to-date reference values
for the various gases (Bickel et al., 2006; Nash et al., 2003; Sartori
et al., 2014): 8.44 €/ton for CO, 2,867.85 €/ton for
NO\textsubscript{X}, 340,969.27 €/ton for PM10, 7,169.62 €/ton for VOC
and 35.85 €/ton for CO\textsubscript{2}eq. The social impacts are in
avoided premature mortality. This result is finally monetized using the
\emph{Statistical Value of Life} for Portugal: €3,055,358/fatality
(Silva et al., 2021). We updated all the monetary reference values of
the literature based on the annual inflation rate in Portugal for
2022\footnote{See
  \href{https://www.ine.pt/xportal/xmain?xpid=INE\&xpgid=ipc}{Statistics
  Portugal tool for inflation rate estimates between years}.}, and our
10-years estimations assumed a discount rate of 5\% and inflation of
3\%. See \protect\hyperlink{research-data}{Research Data} for all the
input values we used.

@@

Shifting from car to cycling in combination with PT can reduce annual
CO\textsubscript{2}eq emissions by 14,000 to 36,000 tons per year. The
10-year socio-environmental benefits account for €125 million to €325
million, depending on the cycling targets.

@@

\hypertarget{acknowledgements}{%
\subsection*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{subsection}{Acknowledgements}

@@

\hypertarget{research-data}{%
\subsection*{Research data}\label{research-data}}
\addcontentsline{toc}{subsection}{Research data}

@@

>>>>

Emission factors are estimated for the following air pollutants: CO,
NO\textsubscript{X}, VOC, and PM\textsubscript{10}. Emission factors of
the main greenhouse gases (GHG) are also estimated: CO\textsubscript{2},
CH\textsubscript{4} and N\textsubscript{2}O, converted in
CO\textsubscript{2}eq by the following relationship\footnote{The weights
  correspond to the Global Warming Potentials (GWP) defined for a
  100-year period by the IPCC in its
  \href{https://www.ipcc.ch/report/ar5/}{5th Assessment Report}.}:
\(EF_{CO_2eq} = EF_{CO_2} + 28\cdot EF_{CH_4} + 265\cdot EF_{N_2O}\).
The CH\textsubscript{4} and N\textsubscript{2}O emission factors do not
vary with travel speed. The PM\textsubscript{10} emission factor does
not vary with trip distance.

@@

The conversion of avoided emissions into avoided welfare loss and
respective monetary valuation was based on the EU Guide to Cost-benefit
Analysis (Sartori et al., 2014) and the best up-to-date reference values
for the various gases (Bickel et al., 2006; Nash et al., 2003; Sartori
et al., 2014): 8.44 €/ton for CO, 2,867.85 €/ton for
NO\textsubscript{X}, 340,969.27 €/ton for PM\textsubscript{10}, 7,169.62
€/ton for VOC and 35.85 €/ton for CO\textsubscript{2}eq. The social
impacts are in avoided premature mortality. This result is finally
monetized using the \emph{Statistical Value of Life} for Portugal:
€3,055,358/fatality (Silva et al., 2021). We updated all the monetary
reference values of the literature based on the annual inflation rate in
Portugal for 2022\footnote{See
  \href{https://www.ine.pt/xportal/xmain?xpid=INE\&xpgid=ipc}{Statistics
  Portugal tool for inflation rate estimates between years}.}, and our
10-years estimations assumed a discount rate of 5\% and inflation of
3\%. See \protect\hyperlink{research-data}{Research Data} for all the
input values we used.

@@

Shifting from car to cycling in combination with PT can reduce annual
CO\textsubscript{2}eq emissions by 11,500 to 28,500 tons per year. These
figures represent a 2.7\% reduction in Lisbon's transport emissions
(Lisboa E-Nova, 2023), a small but important component of wider
transport decarbonization measures. The 10-year socio-environmental
benefits account for €125 million to €325 million, depending on the
cycling targets.

@@

\hypertarget{acknowledgements}{%
\section*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{section}{Acknowledgements}

@@

\hypertarget{research-data}{%
\section*{Research data}\label{research-data}}
\addcontentsline{toc}{section}{Research data}

@@

\leavevmode\vadjust pre{\hypertarget{ref-LisboaENova}{}}%
Lisboa E-Nova. (2023). \emph{{Observatórios Lisboa - Infos Emissões
Gases com Efeito Estufa}}. {Agência de Energia e Ambiente de Lisboa}.
\url{https://observatorios-lisboa.pt/info_emissoes.html}
